Quantum Mechanics Assignment Three

### 2016-01-27 quantum/latex/a3.tex
Add most of hw 3

\textbf{Energy}
\begin{align*}
  <H> &= \bra{1}(\frac{3}{2}\hbar\omega)\ket{1} + \bra{2}(\frac{5}{2}\hbar\omega)\ket{2}\\
  &= 0.36 * \frac{3}{2}\hbar\omega + 0.64 * \frac{5}{2}\hbar\omega\\
  &= 2.08\hbar\omega\\
\end{align*}

@@

\textbf{Normalization constant}
\begin{align*}
  \psi(x,0) &= A\left[\phi_0(x) + 2\phi_1(x) + 2\phi_2(x)\right]\\
  \braket{\psi} &= 1
  A^2\left(1 + 4 + 4\right) &= 1 \rightarrow A = \frac{1}{3}\\
\end{align*}

@@

Standard deviations in QHO systems are interesting because all the energies are evenly spaced. This
value represents how spread-out the states are. It is expected that because the states are close
together for this system, the standard deviation is small.
\end{document}

>>>>

\textbf{Energy}
\begin{align*}
  <H> &= \bra{1}(\frac{3}{2}\hbar\omega)\ket{1} + \bra{2}(\frac{5}{2}\hbar\omega)\ket{2}\\
  &= 0.36 * \frac{3}{2}\hbar\omega + 0.64 * \frac{5}{2}\hbar\omega\\
  &= 2.14\hbar\omega\\
\end{align*}

@@

\textbf{Normalization constant}
\begin{align*}
  \psi(x,0) &= A\left[\phi_0(x) + 2\phi_1(x) + 2\phi_2(x)\right]\\
  \braket{\psi} &= 1\\
  A^2\left(1 + 4 + 4\right) &= 1 \rightarrow A = \frac{1}{3}\\
\end{align*}

@@

Standard deviations in QHO systems are interesting because all the energies are evenly spaced. This
value represents how spread-out the states are. It is expected that because the states are close
together for this system, the standard deviation is small.

\section{9.22: Fictional System}
Here we are given the energy spectrum $E_n = n^3\hbar\omega$ and asked to give some information
about the system: its Hamiltonian and eigenstates in energy representation (matrix form)
and the matrix representation of the operator $A\ket{n} = 3n^2\ket{n+2}$.\\

Because any operator in its own basis is diagonal and because any eigenvector in its own basis is
a unit vector, we represent the Hamiltonian and eigenvectors as so:

\begin{align*}
  H \doteq
  \begin{pmatrix}
    \hbar\omega & 0 & 0 & ...\\
    0 & 8\hbar\omega & 0 & ...\\
    0 & 0 & 27\hbar\omega & ...\\
    ... & ... & ... & ...\\
  \end{pmatrix}
\end{align*}

\begin{align*}
  \ket{1} \doteq
  \begin{pmatrix}
    1\\
    0\\
    0\\
    ...\\
  \end{pmatrix}
  \hspace{1cm}
  \ket{2} \doteq
  \begin{pmatrix}
    0\\
    1\\
    0\\
    ...\\
  \end{pmatrix}
  \hspace{1cm}
  \ket{3} \doteq
  \begin{pmatrix}
    0\\
    0\\
    1\\
    ...\\
  \end{pmatrix}
\end{align*}

To find the matrix representation of A we exploit the fact that in the energy basis eigenstates
are represented by vectors with a single unit entry. Thus the first column of our matrix will only
interact with the n=1 vector, the second column with the n=2 vector, etc. We also know that the
result of our matrix-vector multiplication should produce a vector two energies higher than the
input state. Thus it is easy to see that we will construct a matrix with elements two spaces below
diagonal with values of $3n^2$, where n is the current column.

\begin{align*}
  A \doteq
  \begin{pmatrix}
    0 & 0 & 0 & 0 & 0 & ...  \\
    0 & 0 & 0 & 0 & 0 & ...  \\
    3 & 0 & 0 & 0 & 0 & ...  \\
    0 & 12 & 0 & 0 & 0 & ... \\
    0 & 0 & 27 & 0 & 0 & ... \\
    0 & 0 & 0 & 48 & 0 & ... \\
    ... & ... & ... & ... & ... & ... \\
  \end{pmatrix}
\end{align*}

\section{Paper}
The paper \textit{Quantum ground state and single-phonon control of a mechanical resonator} is a
demonstration that a macroscopic, harmonic-oscillating ``quantum drum'' can be cooled enough to
display behavior normally only predicted for microscopic systems. This is interesting because it
is an easy validation of quantum mechanics. The project entails cooling a resonator until its
heat energy is much smaller than that predicted for its oscillation energy, ensuring that
energy measurements are only of the oscillation. The energy is then measured, and it is shown that
only discrete values are taken on, not a continuum, as predicted by QM.\\

It is interesting to me that the microscopic details of the system need not be taken into account.
It seems like the individual wave functions of each atom in the oscillator can just be summed
into one giant wave function for the oscillator. Maybe that simplification can only be done at
very cold temperatures?\\

\section{10.1: Diagonalize a perturbed Hamiltonian}
We begin with the perturbed Hamiltonian in the $S_z$ basis and convert it into its own perturbed
basis. We then show that the eigenvalues in this basis are identical to those found by redefining
the direction of Z, aka $E_n = \pm \frac{\hbar}{2} \sqrt{(\omega_0 + \omega_1)^2 + \omega_2^2}$.

We begin with the perturbed matrix and diagonalize it:\\

\begin{align*}
  H &\doteq
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    \omega_0 + \omega_1 & \omega_2\\
    \omega_2 & -\omega_0 - \omega_1\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \omega_2\left(\frac{-\omega_2}{\omega_0 + \omega_1}\right)\\
    \omega_2 & -\omega_0 - \omega_1\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
    0 & -\omega_0 - \omega_1 - \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}\\
  &=
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}
  =
  \pm\frac{\hbar}{2}
  \begin{pmatrix}
    -\omega_2 & 0\\
    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
  \end{pmatrix}
\end{align*}

\end{document}
